Cell Decomposition of Rank 2 Quiver Grassmannians

### 2017-06-06 Grassmannian_cells.tex
Inserted section 2

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Torus Actions on Quiver Grassmannians}
\noindent The aim of this section is to define a torus action on quiver Grassmannians which can be used to simplify the calculation of homological invariants in general. If the quiver Grassmannian is smooth, which is for instance the case for exceptional representations by \cite{cr}, it can also be used to stratify the quiver Grassmannians using the results of Bialynicki-Birula. More detailed, let $X$ be a smooth projective variety with a $\C^\ast$-action. For a connecting component of the fixed point set $C\subset X^{\C^\ast}$, we define its attracting set as
\[\mathrm{Att}(C):=\{y\in X\mid \lim_{t\to 0}t.y\in C\}\]

@@

@@

Assume that $U$ is a torus fixed with $M(U)\in X_{\bfi}$. Fix the $w^{\mathrm{th}}$ column $M(U)_w$ and assume that $v_{i_w}\in X_s$. As the matrix representation $M(U)$ is uniquely determined, by the assumption on $d$, it follows that all matrix coefficients corresponding to a basis element $v_i$ such that $v_i\notin X_s$ are zero. This means that $\widehat{M(U)}_w\in X_s$ where $\widehat{M(U)}_w$ is the vector of $U$ corresponding to $M(U)_w$. As the colums are linear independent, this already shows the claim.
\end{proof}
The next step is to generalize this to quiver Grassmannians.
Now let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:i\to j$. We denote by $\Rep(Q)$ the category of $\C$-representations of $Q$. Moreover, let $A_Q\cong \ZZ^{Q_1}$ be the free abelian group and $W_{Q}$ be the free non-abelian group generated by $Q_1$.  
\begin{definition}
The universal abelian covering quiver $\hat Q$ of $Q$ is given by the vertices $\hat Q_0=Q_0\times A_{Q}$ and the arrow set $\hat Q_1=Q_1\times A_{Q}$ where $(\alpha,\chi):(i,\chi)\to (j,\chi+e_\alpha)$ for every $\alpha:i\to j$.

We say that a representation $X\in \Rep(Q)$ can be lifted to $\hat Q$ (resp. $\tilde Q$) if there exists a representation $\hat X\in\Rep(\hat Q)$ (resp. $\tilde X\in\Rep(\tilde Q)$) such that $F_Q\hat X=X$ (resp. $G_Q \tilde X=X$) where $F_Q:\Rep(\hat Q)\to\Rep(Q)$ (resp. $G_Q:\Rep(\tilde Q)\to\Rep(Q)$) is the natural functor.
\end{definition}
Note that we also have a natural functor $H_Q:\Rep(\tilde Q)\to \Rep(\hat Q)$.
Moreover, note that for every connected component of the universal covering quiver of the universal abelian covering quiver is isomorphic to a connected component of the universal covering quiver of the original quiver. The functor $F_Q$ induces a map $F_Q:\ZZ^{\hat Q_1}\to \ZZ^{Q_1}$. We say that a dimension vector $\hat \bfe$ of $\hat Q$ is compatible with $\bfe$ if $F_Q(\hat\bfe)=\bfe$. Moreover, the group $A_Q$ acts on $\hat Q$ via translation inducing an action on $\Rep(\hat Q)$ and on $\ZZ^{\hat Q_1}$. We say two representation (resp. dimension vectors) are equivalent in this case. The analogous observation can also be made for $\tilde Q$.

We choose a map $d:\hat Q_0\to\ZZ$ and fix a representation $X\in\Rep(Q)$ which can be lifted to $\hat Q$. We can consider the decomposition $X_q=\bigoplus_{(q,\chi)\in A_Q} X_{(q,\chi)}$. We define a torus action on each $X_{(q,\chi)}$ via $t.x_{(q,\chi)}=t^{d(q,\chi)}x_{(q,\chi)}$ which can be extended linearly to each $X_q$. For a fixed a subspace $U_q$, we can define the subspace $t.U_q$. In general, this torus action induces no torus action on the Quiver Grassmannians $\Gr_{\bfe}(X)$ as $t.U=(t.U_q)_{q\in Q_0}$ is no subrepresentation of $X$ for every $U\in \Gr_{\bfe}(X)$. Actually, for this the action has to satisfy $X_{\alpha}(t.U_i)\in t.U_j$ for every $\alpha:i\to j$ and every $x\in X_i$. 
\begin{lemma}Fix an integer $c_\alpha\in\ZZ$ for every $\alpha\in Q_1$.
If $X\in\Rep(Q)$ can be lifted, $d:\hat Q_0\to\ZZ$ induces a torus action on $\Gr_\bfe(X)$ if we have $d(j,\chi+e_\alpha)-d(i,\chi)=c_\alpha$ for all $\alpha:i\to j$ and $w\in W_{Q}$.
\end{lemma}
\begin{proof} Let $U\in \Gr_\bfe(X)$. Since $X$ can be lifted, we can write $X_\alpha:X_i\to X_j$ as block matrix consisting of linear maps $X_{(\alpha,\chi)}:X_{(i,\chi)}\to X_{(j,\chi+e_\alpha)}$. Let $u_{(i,\chi)}\in U_{(i,\chi)}$ and $\alpha:i\to j\in Q_1$. Then we have $X_\alpha(u_{(i,\chi)})\in U_{(j,\chi+e_\alpha)}$ because $X$ can be lifted to the universal cover. Then it follows
\[X_\alpha(t.u_{(i,\chi)})=X_\alpha(t^{d(i,\chi)}u_{(i,\chi)})=t^{d(i,\chi)}X_\alpha(u_{(i,\chi)})=t^{d(i,\chi)}u_{(j,\chi+e_\alpha)}=t^{-c_\alpha}t.u_{(j,\chi+e_\alpha)}\]
for some $u_{(j,\chi+e_\alpha)}\in U_{(j,\chi+e_\alpha)}$.
Now we have \[\bigoplus_{\chi\in A_Q} t.U_{(j,\chi)}=\bigoplus_{\chi\in A_Q}t^{c(j,\chi)}t.U_{(j,\chi)}\]
for every $t\in\C^\ast$ if $c(j,\chi)$ is constant along $(j,\chi)\in \{j\}\times A_Q$. This yields  
\[t.U_j=\bigoplus_{\chi\in A_Q} t.U_{(j,\chi)}=\bigoplus_{\chi\in A_Q}t^{-c_\alpha}t.U_{(j,\chi)}\]
for every $t\in \C^\ast$.
\end{proof}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Terminology and Preliminary results}
\subsection{Covering theory}
Let $Q$ be an acyclic quiver with vertices $Q_0$ and arrows $Q_1$ which we denote by $\alpha:i\to j$. We denote by $\Rep(Q)$ the category of $\C$-representations of $Q$. Moreover, let $A_Q\cong \ZZ^{Q_1}$ be the free abelian group and $W_{Q}$ be the free non-abelian group generated by $Q_1$.  
\begin{definition}
The universal abelian covering quiver $\hat Q$ of $Q$ is given by the vertices $\hat Q_0=Q_0\times A_{Q}$ and the arrow set $\hat Q_1=Q_1\times A_{Q}$ where $(\alpha,\chi):(i,\chi)\to (j,\chi+e_\alpha)$ for every $\alpha:i\to j$.

We say that a representation $X\in \Rep(Q)$ can be lifted to $\hat Q$ (resp. $\tilde Q$) if there exists a representation $\hat X\in\Rep(\hat Q)$ (resp. $\tilde X\in\Rep(\tilde Q)$) such that $F_Q\hat X=X$ (resp. $G_Q \tilde X=X$) where $F_Q:\Rep(\hat Q)\to\Rep(Q)$ (resp. $G_Q:\Rep(\tilde Q)\to\Rep(Q)$) is the natural functor.
\end{definition}
Note that we also have a natural functor $H_Q:\Rep(\tilde Q)\to \Rep(\hat Q)$.
Moreover, note that every connected component of the universal covering quiver of the universal abelian covering quiver is isomorphic to a connected component of the universal covering quiver of the original quiver. The functor $F_Q$ induces a map $F_Q:\ZZ^{\hat Q_1}\to \ZZ^{Q_1}$. We say that a dimension vector $\hat \bfe$ of $\hat Q$ is compatible with $\bfe$ if $F_Q(\hat\bfe)=\bfe$. The group $A_Q$ acts on $\hat Q$ via translation inducing an action on $\Rep(\hat Q)$ and on $\ZZ^{\hat Q_1}$. We say two representation (resp. dimension vectors) are equivalent in this case. The analogous observation can also be made for $\tilde Q$.

\subsection{Bialynicki-Birula decomposition}
The aim of this section is to define a torus action on quiver Grassmannians which can be used to simplify the calculation of homological invariants in general. If the quiver Grassmannian is smooth, which is for instance the case for exceptional representations by \cite{cr}, it can also be used to stratify the quiver Grassmannians using the results of Bialynicki-Birula. More detailed, let $X$ be a smooth projective variety with a $\C^\ast$-action. For a connecting component of the fixed point set $C\subset X^{\C^\ast}$, we define its attracting set as
\[\mathrm{Att}(C):=\{y\in X\mid \lim_{t\to 0}t.y\in C\}\]

@@

\section{Torus Actions on Quiver Grassmannians}
\noindent

@@

Assume that $U$ is a torus fixed with $M(U)\in X_{\bfi}$. Fix the $w^{\mathrm{th}}$ column $M(U)_w$ and assume that $v_{i_w}\in X_s$. As the matrix representation $M(U)$ is uniquely determined, by the assumption on $d$, it follows that all matrix coefficients corresponding to a basis element $v_i$ such that $v_i\notin X_s$ are zero. This means that $\widehat{M(U)}_w\in X_s$ where $\widehat{M(U)}_w$ is the vector of $U$ corresponding to $M(U)_w$. As the colums are linear independent, this already shows the claim.
\end{proof}
The next step is to generalize this to quiver Grassmannians.

Let $Q$ be an acyclic quiver. We choose a map $d:\hat Q_0\to\ZZ$ and fix a representation $X\in\Rep(Q)$ which can be lifted to $\hat Q$. We can consider the decomposition $X_q=\bigoplus_{(q,\chi)\in A_Q} X_{(q,\chi)}$. We define a torus action on each $X_{(q,\chi)}$ via $t.x_{(q,\chi)}=t^{d(q,\chi)}x_{(q,\chi)}$ which can be extended linearly to each $X_q$. For a fixed a subspace $U_q$, we can define the subspace $t.U_q$. In general, this torus action induces no torus action on the Quiver Grassmannians $\Gr_{\bfe}(X)$ as $t.U=(t.U_q)_{q\in Q_0}$ is no subrepresentation of $X$ for every $U\in \Gr_{\bfe}(X)$. Actually, such an action has to satisfy $X_{\alpha}(t.U_i)\in t.U_j$ for every $\alpha:i\to j$. 
\begin{lemma}Fix an integer $c_\alpha\in\ZZ$ for every $\alpha\in Q_1$.
If $X\in\Rep(Q)$ can be lifted, $d:\hat Q_0\to\ZZ$ induces a torus action on $\Gr_\bfe(X)$ if we have $d(j,\chi+e_\alpha)-d(i,\chi)=c_\alpha$ for all $\alpha:i\to j$ and $\chi\in A_{Q}$.
\end{lemma}
\begin{proof} Let $U\in \Gr_\bfe(X)$. Since $X$ can be lifted, we can write $X_\alpha:X_i\to X_j$ as block matrix consisting of linear maps $X_{(\alpha,\chi)}:X_{(i,\chi)}\to X_{(j,\chi+e_\alpha)}$. Let $u_{(i,\chi)}\in U_{(i,\chi)}$ and $\alpha:i\to j\in Q_1$. Then we have $X_\alpha(u_{(i,\chi)})\in U_{(j,\chi+e_\alpha)}$ because $X$ can be lifted to the universal cover. Then it follows
\[X_\alpha(t.u_{(i,\chi)})=X_\alpha(t^{d(i,\chi)}u_{(i,\chi)})=t^{d(i,\chi)}X_\alpha(u_{(i,\chi)})=t^{d(i,\chi)}u_{(j,\chi+e_\alpha)}=t^{-c_\alpha}t.u_{(j,\chi+e_\alpha)}\]
for some $u_{(j,\chi+e_\alpha)}\in U_{(j,\chi+e_\alpha)}$.
Now we have \[\bigoplus_{\chi\in A_Q} t.U_{(j,\chi)}=\bigoplus_{\chi\in A_Q}t^{c(j,\chi)}t.U_{(j,\chi)}\]
for every $t\in\C^\ast$ if $c(j,\chi)$ is constant along $(j,\chi)\in \{j\}\times A_Q$. This yields  
\[t.U_j=\bigoplus_{\chi\in A_Q} t.U_{(j,\chi)}=\bigoplus_{\chi\in A_Q}t^{-c_\alpha}t.U_{(j,\chi)}\]
for every $t\in \C^\ast$.
\end{proof}
